```{r}
install.packages("remotes")
remotes::install_github("kevinblighe/CorLevelPlot")
```

```{r}
library(CorLevelPlot)
library(tidyverse)
library(edgeR)
library(WGCNA)
library(DESeq2)
library(GEOquery)
library(tidyverse)
library(gridExtra)
```

# Loading the MetaData (data_info)

```{r}
sample_info <- as.data.frame(read_excel("/home/yassine/Downloads/Repro_omics_sample_information_06_27_2023.xlsx"))

```

```{r}
data_info <- sample_info[, -4] %>% # Remove the fourth column
  filter(!is.na(sample_info[[5]])) %>% #Keep rows with non-NA entries in the fifth column (tissu)
  dplyr::rename(tissu = sample)%>% # Rename the 'sample' column to 'tissu'
  filter(!is.na(heiferID))%>% #Remove rows with NA entry in the 'heiferID' column
  mutate(month = as.integer(gsub("-month", "", month)))#Convert character entries in the 'month' column to integers

colnames(data_info)[4] <- "sample"
data_info$sample_name <- paste0(data_info$heiferID, "_", data_info$tissu) #To avoid having duplicate samples later 
```

```{r}
print(data_info)
```

```{r}
data_info$sample<- as.factor(data_info$sample)
data_info$month<- as.factor(data_info$month)
data_info$tissu<- as.factor(data_info$tissu)
data_info$heiferID<- as.factor(data_info$heiferID)

```

#Loading the count data 
```{r}
files<-list.files("/mnt/storage/lab_folder/repro_omics/counting/", 
                  recursive=T,
                  pattern="count", 
                  full.names = TRUE)

files<-files[grep("summary", files, invert = TRUE)]
files<-files[grep(".count1", files, invert = TRUE)]
files<-files[grep(".count2", files, invert = TRUE)]
files<-files[grep(".sh", files, invert = TRUE)]

count_data<-data.frame(matrix(nrow=27607))
for (n in 1:132) {
  count<-read.delim(files[n],
                    header=TRUE, 
                    sep= "\t", 
                    stringsAsFactors = FALSE,
                    comment.char= "#")
  
  count<-count[,c(1,7)]
  count_data<-cbind(count_data,count)
}
rownames(count_data)<-count_data[,2]
count_data<-count_data[,seq(from = 3, 
                            to = 265,
                            by = 2)]
colnames(count_data)<- substr(colnames(count_data),57, 59)
```

```{r}
colnames(count_data)<-data_info$sample_name
```

# Removing genes that have no reads in any samples -> count_data_a

# Adding the annotation from "taurus_gene_ensembl" to the filtred gene counts  ->  count_data_annotated

# Keep only genes transscripts that are 'protein_coding', 'lncRNA' or 'pseudogene' -b count_data_a

```{r}
count_data_a<-count_data[rowSums(count_data)>0,] #filtering genes that have no reads 

# Adds the annotation from "taurus_gene_ensembl" to the filtred gene counts  
count_data_annotated<-merge(count_data_a,
                            annotation.ensembl.symbol,
                            by.x="row.names",
                            by.y="ensembl_gene_id",
                            all.x=TRUE,
                            all.y=FALSE)
```


```{r}
count_data_annotated<-count_data_annotated[count_data_annotated$gene_biotype %in% c('protein_coding', 'lncRNA','pseudogene'),]

count_data_b<-count_data_annotated[,c(2:133)]

rownames(count_data_b)<-count_data_annotated$Row.names
```

# CTF_normalizing the count data -> CTF_normalized
```{r}
lib_size <- base::colSums(count_data_b)

#TMM normalization factors
norm_factors <- calcNormFactors(object = count_data_b,
                                  lib.size = lib_size,
                                  method = "TMM")
#divide with norm factors here
CTF_normalized <- sweep(count_data_b, 2, norm_factors, "/")

```

# Transforming the entries of CTF_normalized data-frame with the asinh function

```{r}
CTF_normalized_asinh <- asinh(CTF_normalized)
```

#  WGCNA default filter 

```{r}
gsg <- goodSamplesGenes(t(CTF_normalized_asinh))
summary(gsg)
gsg$allOK
```

```{r}
table(gsg$goodGenes)
table(gsg$goodSamples)
```
# Outlier detection and removal before co-expression analysis 

# - hierarchical clustering - method 1

```{r}
htree <- hclust(dist(t(CTF_normalized_asinh)), method = "average")
plot(htree)
```

# - PCA  - method 2

```{r}
pca <- prcomp(t(CTF_normalized_asinh))
pca.dat <- pca$x

pca.var <- pca$sdev^2
pca.var.percent <- round(pca.var/sum(pca.var)*100, digits = 2)

pca.dat <- as.data.frame(pca.dat)

ggplot(pca.dat, aes(PC1, PC2)) +
  geom_point() +
  geom_text(label = rownames(pca.dat)) +
  labs(x = paste0('PC1: ', pca.var.percent[1], ' %'),
       y = paste0('PC2: ', pca.var.percent[2], ' %'))
```

# Remove the samples detected as outliers
```{r}
# exclude outlier samples
samples.to.be.excluded <- c('H392B_endometrium','J148B_hypothalamus','J133B_hypothalamus','J134B_hypothalamus','H273B_hypothalamus','H378B_hypothalamus')
CTF_normalized_asinh <- CTF_normalized_asinh[,!(colnames(CTF_normalized_asinh) %in% samples.to.be.excluded)]
```
# Network Construction 

# First step : Choose a set of soft-thresholding powers
```{r}
power <- c(c(1:10), seq(from = 1, to = 30, by = 1))

# Call the network topology analysis function
sft <- pickSoftThreshold(t(CTF_normalized_asinh),
                  powerVector = power,
                  networkType = "signed",
                  verbose = 5)

sft.data <- sft$fitIndices
```

# visualization to pick power

```{r}

a1 <- ggplot(sft.data, aes(Power, SFT.R.sq, label = Power)) +
  geom_point() +
  geom_text(nudge_y = 0.1) +
  geom_hline(yintercept = 0.8, color = 'red') +
  labs(x = 'Power', y = 'Scale free topology model fit, signed R^2') +
  theme_classic()


a2 <- ggplot(sft.data, aes(Power, mean.k., label = Power)) +
  geom_point() +
  geom_text(nudge_y = 0.1) +
  labs(x = 'Power', y = 'Mean Connectivity') +
  theme_classic()
  

grid.arrange(a1, a2, nrow = 2)
```

```{r}
CTF_normalized_asinh_ <-t(CTF_normalized_asinh)
```


```{r}
# convert matrix to numeric
CTF_normalized_asinh_[] <- sapply(CTF_normalized_asinh_, as.numeric)

soft_power <- 17
temp_cor <- cor
cor <- WGCNA::cor


# memory estimate w.r.t blocksize
bwnet <- blockwiseModules(CTF_normalized_asinh_,
                 maxBlockSize = 22600,
                 TOMType = "signed",
                 power = soft_power,
                 mergeCutHeight = 0.25,
                 numericLabels = FALSE,
                 randomSeed = 1234,
                 verbose = 3,
                 nThreads = 10)
```


```{r}
cor <- temp_cor
```

# Formation of gene modules 

```{r}
module_eigengenes <- bwnet$MEs

# get number of genes for each module
table(bwnet$colors)
```

# Cluster dendogram 

```{r}

# Plot the dendrogram and the module colors before and after merging underneath
plotDendroAndColors(bwnet$dendrograms[[1]], cbind(bwnet$unmergedColors, bwnet$colors),
                    c("unmerged", "merged"),
                    dendroLabels = FALSE,
                    addGuide = TRUE,
                    hang= 0.03,
                    guideHang = 0.05)
```

# Pairwise pearson correlation matrix 

```{r}
colData<-data_info
row.names(colData)<-data_info$sample_name
# binarize categorical variables
colData$month <- factor(colData$month, levels = c("6", "9", "19"))
colData$tissu <- factor(colData$tissu, levels = c("endometrium", "hypothalamus", "pwbc", "pituitary"))

month.out <- binarizeCategoricalColumns(colData$month,
                           includePairwise = TRUE,
                           includeLevelVsAll = FALSE,
                           minCount = 1)

tissu.out <- binarizeCategoricalColumns(colData$tissu,
                           includePairwise = TRUE,
                           includeLevelVsAll = FALSE,
                           minCount = 1)

traits <- cbind(month.out, tissu.out)

# Define numbers of genes and samples
nsamples <- nrow(CTF_normalized_asinh_)
ngenes <- ncol(CTF_normalized_asinh_)


module.trait.corr <- cor(module_eigengenes,
                         traits,
                         use = 'p')

module.trait.corr.pvals <- corPvalueStudent(module.trait.corr,
                                              nsamples)
```


# Heat map visualisation of the Pearson correlation matrix
```{r}
row.names(traits)<-row.names(module_eigengenes)

# visualize module-trait association as a heatmap
heatmap.data <- merge(module_eigengenes,
                      traits,
                      by = 'row.names')
heatmap.data <- heatmap.data %>% 
  column_to_rownames(var = 'Row.names')
```

```{r}
names(heatmap.data)
```


```{r}
names(heatmap.data)[20:28]<-c("9.vs.6",
                                 "19.vs.6"                    
                                 ,"19.vs.9",
                                 "hypoth.vs.endomet",
                                 "pwbc.vs.endomet",
                                 "pituit.vs.endomet" ,  
                                 "pwbc.vs.hypoth",
                                 "pituit.vs.hypoth",
                                 "pituit.vs.pwbc")
```

```{r}
CorLevelPlot(heatmap.data,
             x = names(heatmap.data)[20:28],
             y = names(heatmap.data)[1:19],
             col = c("blue1", "skyblue", "white", "pink", "red"))
```

# Intramodular analysis: Identification of driver genes 
```{r}
heatmap.data
```

```{r}

# Calculate the module membership and the associated p-values

# The module membership quantifies the similarity of all genes on the array to every module.

module.membership.measure <- cor(module_eigengenes,
                                   CTF_normalized_asinh_,
                                   use = 'p')
module.membership.measure.pvals <- corPvalueStudent(module.membership.measure,
                                                    nsamples)
```


```{r}
View(module.membership.measure)
```

```{r}
# Calculate the gene significance and associated p-values

hypothalamus.vs.endometrium_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.hypothalamus.vs.endometrium,
                                    use = 'p')
pwbc.vs.endometrium_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pwbc.vs.endometrium,
                                    use = 'p')
pituitary.vs.endometrium_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pituitary.vs.endometrium,
                                    use = 'p')
pwbc.vs.hypothalamus_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pwbc.vs.hypothalamus,
                                    use = 'p')
pituitary.vs.hypothalamus_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pituitary.vs.hypothalamus,
                                    use = 'p')
pituitary.vs.pwbc_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pituitary.vs.pwbc,
                                    use = 'p')

hypothalamus.vs.endometrium_gene.signf.pvals <- corPvalueStudent(hypothalamus.vs.endometrium_gene.signf.corr,nsamples)

pwbc.vs.endometrium_gene.signf.pvals <- corPvalueStudent(pwbc.vs.endometrium_gene.signf.corr,nsamples)

pituitary.vs.endometrium_gene.signf.pvals <- corPvalueStudent(pituitary.vs.endometrium_gene.signf.corr,nsamples)

pwbc.vs.hypothalamus_gene.signf.pvals <- corPvalueStudent(pwbc.vs.hypothalamus_gene.signf.corr,nsamples)

pituitary.vs.hypothalamus_gene.signf.pvals <- corPvalueStudent(pituitary.vs.hypothalamus_gene.signf.corr,nsamples)

pituitary.vs.pwbc_gene.signf.pvals <- corPvalueStudent(pituitary.vs.pwbc_gene.signf.corr,nsamples)

```
```{r}
n = nrow(hypothalamus.vs.endometrium_gene.signf.pvals)
alpha = 0.01
adjusted_p_value = alpha / n 

```

```{r}
adjusted_p_value
```

# Signficantly correlated genes hypothalamus.vs.endometrium
```{r}
hypothalamus.vs.endometrium_gene.signf.pvals<-data.frame(hypothalamus.vs.endometrium_gene.signf.pvals)
colnames(hypothalamus.vs.endometrium_gene.signf.pvals)<- c("p_value")

hypothalamus.vs.endometrium_gene.signf.pvals$geneID <- row.names(hypothalamus.vs.endometrium_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than adjusted_p_value
filtered_hypothalamus.vs.endometrium_gene.signf.pvals <- hypothalamus.vs.endometrium_gene.signf.pvals[hypothalamus.vs.endometrium_gene.signf.pvals$p_value <= adjusted_p_value,]
filtered_hypothalamus.vs.endometrium_gene.signf.pvals <- na.omit(filtered_hypothalamus.vs.endometrium_gene.signf.pvals)
# Print the filtered data frame
print(filtered_hypothalamus.vs.endometrium_gene.signf.pvals)

```


# Signficantly correlated genes pwbc.vs.endometrium


```{r}
pwbc.vs.endometrium_gene.signf.pvals<-data.frame(pwbc.vs.endometrium_gene.signf.pvals)
colnames(pwbc.vs.endometrium_gene.signf.pvals)<- c("p_value")

pwbc.vs.endometrium_gene.signf.pvals$geneID <- row.names(pwbc.vs.endometrium_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than adjusted_p_value
filtered_pwbc.vs.endometrium_gene.signf.pvals <- pwbc.vs.endometrium_gene.signf.pvals[pwbc.vs.endometrium_gene.signf.pvals$p_value <= adjusted_p_value,]
pwbc.vs.endometrium_gene.signf.pvals <- na.omit(filtered_pwbc.vs.endometrium_gene.signf.pvals)

# Print the filtered data frame
print(filtered_pwbc.vs.endometrium_gene.signf.pvals)

```
# Signficantly correlated genes pituitary.vs.endometrium

```{r}

pituitary.vs.endometrium_gene.signf.pvals<-data.frame(pituitary.vs.endometrium_gene.signf.pvals)
colnames(pituitary.vs.endometrium_gene.signf.pvals)<- c("p_value")

pituitary.vs.endometrium_gene.signf.pvals$geneID <- row.names(pituitary.vs.endometrium_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than adjusted_p_value
filtered_pituitary.vs.endometrium_gene.signf.pvals <- pituitary.vs.endometrium_gene.signf.pvals[pituitary.vs.endometrium_gene.signf.pvals$p_value <= adjusted_p_value,]
pituitary.vs.endometrium_gene.signf.pvals <- na.omit(filtered_pituitary.vs.endometrium_gene.signf.pvals)

# Print the filtered data frame
print(pituitary.vs.endometrium_gene.signf.pvals)
```

#  Signficantly correlated genes pwbc.vs.hypothalamus
```{r}
pwbc.vs.hypothalamus_gene.signf.pvals<-data.frame(pwbc.vs.hypothalamus_gene.signf.pvals)
colnames(pwbc.vs.hypothalamus_gene.signf.pvals)<- c("p_value")

pwbc.vs.hypothalamus_gene.signf.pvals$geneID <- row.names(pwbc.vs.hypothalamus_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than 0.01
filtered_pwbc.vs.hypothalamus_gene.signf.pvals <- pwbc.vs.hypothalamus_gene.signf.pvals[pwbc.vs.hypothalamus_gene.signf.pvals$p_value <= adjusted_p_value,]
pwbc.vs.hypothalamus_gene.signf.pvals <- na.omit(filtered_pwbc.vs.hypothalamus_gene.signf.pvals)

# Print the filtered data frame
print(filtered_pwbc.vs.hypothalamus_gene.signf.pvals)
```

#  Signficantly correlated genes pituitary.vs.hypothalamus

```{r}
pituitary.vs.hypothalamus_gene.signf.pvals<-data.frame(pituitary.vs.hypothalamus_gene.signf.pvals)
colnames(pituitary.vs.hypothalamus_gene.signf.pvals)<- c("p_value")

pituitary.vs.hypothalamus_gene.signf.pvals$geneID <- row.names(pituitary.vs.hypothalamus_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than 0.01
filtered_pituitary.vs.hypothalamus_gene.signf.pvals <- pituitary.vs.hypothalamus_gene.signf.pvals[pituitary.vs.hypothalamus_gene.signf.pvals$p_value <= adjusted_p_value,]
pituitary.vs.hypothalamus_gene.signf.pvals <- na.omit(filtered_pituitary.vs.hypothalamus_gene.signf.pvals)

# Print the filtered data frame
print(filtered_pituitary.vs.hypothalamus_gene.signf.pvals)
```


# Signficantly correlated genes pituitary.vs.pwbc : 
```{r}
pituitary.vs.pwbc_gene.signf.pvals<-data.frame(pituitary.vs.pwbc_gene.signf.pvals)
colnames(pituitary.vs.pwbc_gene.signf.pvals)<- c("p_value")

pituitary.vs.pwbc_gene.signf.pvals$geneID <- row.names(pituitary.vs.pwbc_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than 0.01
filtered_pituitary.vs.pwbc_gene.signf.pvals <- pituitary.vs.pwbc_gene.signf.pvals[pituitary.vs.pwbc_gene.signf.pvals$p_value <= adjusted_p_value,]
pituitary.vs.pwbc_gene.signf.pvals <- na.omit(filtered_pituitary.vs.pwbc_gene.signf.pvals)

# Print the filtered data frame
print(filtered_pituitary.vs.pwbc_gene.signf.pvals)
```

```{r}
# Combine the geneID columns into a single vector
all_de.gene_ids <- c(
  pituitary.vs.hypothalamus_gene.signf.pvals$geneID,
  pwbc.vs.hypothalamus_gene.signf.pvals$geneID,
  pituitary.vs.endometrium_gene.signf.pvals$geneID,
  pwbc.vs.endometrium_gene.signf.pvals$geneID,
  hypothalamus.vs.endometrium_gene.signf.pvals$geneID,
  pituitary.vs.pwbc_gene.signf.pvals$geneID
)
```

```{r}
all_de.gene_ids
```

```{r}
# Remove duplicates from the geneID vector
unique_de.gene_ids <- unique(all_de.gene_ids)
```

```{r}
measured_genes<-colnames(CTF_normalized_asinh_)
```

```{r}
length(unique_de.gene_ids)
