# Read the count data

```{r}
files<-list.files("/mnt/storage/lab_folder/repro_omics/counting/", 
                  recursive=T,
                  pattern="count", 
                  full.names = TRUE)

files<-files[grep("summary", files, invert = TRUE)]
files<-files[grep(".count1", files, invert = TRUE)]
files<-files[grep(".count2", files, invert = TRUE)]
files<-files[grep(".sh", files, invert = TRUE)]

#length(files)

```
```{r}
count_data<-data.frame(matrix(nrow=27607))
for (n in 1:132) {
  count<-read.delim(files[n],
                    header=TRUE, 
                    sep= "\t", 
                    stringsAsFactors = FALSE,
                    comment.char= "#")
  
  count<-count[,c(1,7)]
  count_data<-cbind(count_data,count)
}
rownames(count_data)<-count_data[,2]
count_data<-count_data[,seq(from = 3, 
                            to = 265,
                            by = 2)]
colnames(count_data)<- substr(colnames(count_data),57, 59)
```

# Import Sample Information

```{r}
sample_info <- as.data.frame(read_excel("/home/yassine/Downloads/Repro_omics_sample_information_06_27_2023.xlsx"))
```

# Wrangling and data exploration

```{r}
sample_info_ <- sample_info[, -c(4,5)]

sample_info_ <- sample_info_[!is.na(sample_info_$heiferID), ]

sample_info_ <- sample_info_ %>%
  mutate(month = as.integer(gsub("-month", "", month))) %>%
  dplyr::rename(tissue = sample)

# creation of the data_info for the co-expression analysis

```{r}
data_info <- sample_info[, -4] %>% # Remove the fourth column
  filter(!is.na(sample_info[[5]])) %>% #Keep rows with non-NA entries in the fifth column (tissu)
  dplyr::rename(tissu = sample)%>% # Rename the 'sample' column to 'tissu'
  filter(!is.na(heiferID))%>% #Remove rows with NA entry in the 'heiferID' column
  mutate(month = as.integer(gsub("-month", "", month)))#Convert character entries in the 'month' column to integers

colnames(data_info)[4] <- "sample"
```

# Summary dataframe for all samples (counts of unique tissues and their names per month)

```{r}

month_tissue_summary_ <- data_info %>%
  group_by(month) %>%
  summarise(
    tissues = paste(unique(tissu), collapse = ", "),
    distinct_samples = n_distinct(heiferID),
    Heifer_IDs = paste(unique(heiferID), collapse = ", ")
  )
# Print the modified summary dataframe
print(month_tissue_summary_)
```
# remove samples with low reads generated 
```{r}
# List of sample IDs to remove
samples_to_remove <- c(8, 16, 56, 57, 59, 102)

# Remove rows with specified sample IDs
filtred_data_info <- data_info %>%
  filter(!(sample %in% samples_to_remove))

```
```{r}
# Extract the rows of data_info that have sample column have one of these values
samples_removed <- data_info[(data_info$sample %in% samples_to_remove), ]

```
# Summary dataframe of the filtred dataset (counts of unique tissues and their names per month)

```{r}
# Calculate the distinct samples per category (month and tissu)
month_tissue_summary_ <- filtred_data_info %>%
  group_by(month, tissu) %>%
  summarise(
    distinct_samples = n_distinct(sample),
    sample_names = paste(unique(sample), collapse = ", ")
  )

# Print the modified summary dataframe
print(month_tissue_summary_)

```
```{r}
data_info$sample<- as.factor(data_info$sample)
data_info$month<- as.factor(data_info$month)
data_info$tissu<- as.factor(data_info$tissu)
data_info$heiferID<- as.factor(data_info$heiferID)

colnames(count_data)<-data_info$heiferID

count_data_export<-data.frame(gene_id=rownames(count_data),count_data)

write_delim(count_data_export,
            file = "/home/yassine/Downloads/unfiltered_count_data.txt", 
            delim = "\t",
            quote =  "none")

#system("bzip2 "/home/yassine/Downloads/unfiltered_count_data.txt")


write_delim(data_info,
            file = "/home/yassine/Downloads/unfiltered_data_info.txt",
            delim = "\t",
            quote =  "none")

#system("bzip2 "/home/yassine/Downloads/unfiltered_data_info.txt)

```
# Filter the count data (keep only : protein_coding, lncRNA, pseudogene)

```{r}
count_data_a<-count_data[rowSums(count_data)>0,] #filtering genes that have no reads 

# Adds the annotation from "btaurus_gene_ensembl" to the filtred gene counts  
count_data_annotated<-merge(count_data_a,
                            annotation.ensembl.symbol,
                            by.x="row.names",
                            by.y="ensembl_gene_id",
                            all.x=TRUE,
                            all.y=FALSE)
```


```{r}
# Count_data_annotated_length contains the read counts and the transcript_lengths for the genes that generates one of these RNA types 'protein_coding', 'lncRNA','pseudogene' the count_data_annotated data frame.
count_data_annotated<-count_data_annotated[count_data_annotated$gene_biotype %in% c('protein_coding', 'lncRNA','pseudogene'),]
count_data_annotated_length<-gene.length[gene.length$ensembl_gene_id %in% count_data_annotated$Row.names, 2]
count_data_b<-count_data_annotated[,c(2:123)]
rownames(count_data_b)<-count_data_annotated$Row.names
```

```{r}
write_delim(count_data_b,
            file = "/home/yassine/Downloads/filtered_count_data.txt",
            delim = "\t",
            quote =  "none")

#system("bzip2 "/home/yassine/Downloads/filtered_count_data.txt")
```
