# Intramodular analysis: Identification of driver genes 

```{r}

# Calculate the module membership and the associated p-values

# The module membership quantifies the similarity of all genes on the array to every module.

module.membership.measure <- cor(module_eigengenes,
                                   CTF_normalized_asinh_,
                                   use = 'p')
module.membership.measure.pvals <- corPvalueStudent(module.membership.measure,
                                                    nsamples)
```

```{r}
# Calculate the gene significance and associated p-values

hypothalamus.vs.endometrium_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.hypothalamus.vs.endometrium,
                                    use = 'p')
pwbc.vs.endometrium_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pwbc.vs.endometrium,
                                    use = 'p')
pituitary.vs.endometrium_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pituitary.vs.endometrium,
                                    use = 'p')
pwbc.vs.hypothalamus_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pwbc.vs.hypothalamus,
                                    use = 'p')
pituitary.vs.hypothalamus_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pituitary.vs.hypothalamus,
                                    use = 'p')
pituitary.vs.pwbc_gene.signf.corr <- cor(CTF_normalized_asinh_,
                                    traits$data.pituitary.vs.pwbc,
                                    use = 'p')

hypothalamus.vs.endometrium_gene.signf.pvals <- corPvalueStudent(hypothalamus.vs.endometrium_gene.signf.corr,nsamples)

pwbc.vs.endometrium_gene.signf.pvals <- corPvalueStudent(pwbc.vs.endometrium_gene.signf.corr,nsamples)

pituitary.vs.endometrium_gene.signf.pvals <- corPvalueStudent(pituitary.vs.endometrium_gene.signf.corr,nsamples)

pwbc.vs.hypothalamus_gene.signf.pvals <- corPvalueStudent(pwbc.vs.hypothalamus_gene.signf.corr,nsamples)

pituitary.vs.hypothalamus_gene.signf.pvals <- corPvalueStudent(pituitary.vs.hypothalamus_gene.signf.corr,nsamples)

pituitary.vs.pwbc_gene.signf.pvals <- corPvalueStudent(pituitary.vs.pwbc_gene.signf.corr,nsamples)

```

```{r}
n = nrow(hypothalamus.vs.endometrium_gene.signf.pvals)
alpha = 0.01
adjusted_p_value = alpha / n 

```


# Signficantly correlated genes hypothalamus.vs.endometrium

```{r}
hypothalamus.vs.endometrium_gene.signf.pvals<-data.frame(hypothalamus.vs.endometrium_gene.signf.pvals)
colnames(hypothalamus.vs.endometrium_gene.signf.pvals)<- c("p_value")

hypothalamus.vs.endometrium_gene.signf.pvals$geneID <- row.names(hypothalamus.vs.endometrium_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than adjusted_p_value
filtered_hypothalamus.vs.endometrium_gene.signf.pvals <- hypothalamus.vs.endometrium_gene.signf.pvals[hypothalamus.vs.endometrium_gene.signf.pvals$p_value <= adjusted_p_value,]
filtered_hypothalamus.vs.endometrium_gene.signf.pvals <- na.omit(filtered_hypothalamus.vs.endometrium_gene.signf.pvals)
# Print the filtered data frame
print(filtered_hypothalamus.vs.endometrium_gene.signf.pvals)

```


# Signficantly correlated genes pwbc.vs.endometrium


```{r}
pwbc.vs.endometrium_gene.signf.pvals<-data.frame(pwbc.vs.endometrium_gene.signf.pvals)
colnames(pwbc.vs.endometrium_gene.signf.pvals)<- c("p_value")

pwbc.vs.endometrium_gene.signf.pvals$geneID <- row.names(pwbc.vs.endometrium_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than adjusted_p_value
filtered_pwbc.vs.endometrium_gene.signf.pvals <- pwbc.vs.endometrium_gene.signf.pvals[pwbc.vs.endometrium_gene.signf.pvals$p_value <= adjusted_p_value,]
pwbc.vs.endometrium_gene.signf.pvals <- na.omit(filtered_pwbc.vs.endometrium_gene.signf.pvals)

# Print the filtered data frame
print(filtered_pwbc.vs.endometrium_gene.signf.pvals)

```
# Signficantly correlated genes pituitary.vs.endometrium

```{r}

pituitary.vs.endometrium_gene.signf.pvals<-data.frame(pituitary.vs.endometrium_gene.signf.pvals)
colnames(pituitary.vs.endometrium_gene.signf.pvals)<- c("p_value")

pituitary.vs.endometrium_gene.signf.pvals$geneID <- row.names(pituitary.vs.endometrium_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than adjusted_p_value
filtered_pituitary.vs.endometrium_gene.signf.pvals <- pituitary.vs.endometrium_gene.signf.pvals[pituitary.vs.endometrium_gene.signf.pvals$p_value <= adjusted_p_value,]
pituitary.vs.endometrium_gene.signf.pvals <- na.omit(filtered_pituitary.vs.endometrium_gene.signf.pvals)

# Print the filtered data frame
print(pituitary.vs.endometrium_gene.signf.pvals)
```

#  Signficantly correlated genes pwbc.vs.hypothalamus
```{r}
pwbc.vs.hypothalamus_gene.signf.pvals<-data.frame(pwbc.vs.hypothalamus_gene.signf.pvals)
colnames(pwbc.vs.hypothalamus_gene.signf.pvals)<- c("p_value")

pwbc.vs.hypothalamus_gene.signf.pvals$geneID <- row.names(pwbc.vs.hypothalamus_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than 0.01
filtered_pwbc.vs.hypothalamus_gene.signf.pvals <- pwbc.vs.hypothalamus_gene.signf.pvals[pwbc.vs.hypothalamus_gene.signf.pvals$p_value <= adjusted_p_value,]
pwbc.vs.hypothalamus_gene.signf.pvals <- na.omit(filtered_pwbc.vs.hypothalamus_gene.signf.pvals)

# Print the filtered data frame
print(filtered_pwbc.vs.hypothalamus_gene.signf.pvals)
```

#  Signficantly correlated genes pituitary.vs.hypothalamus

```{r}
pituitary.vs.hypothalamus_gene.signf.pvals<-data.frame(pituitary.vs.hypothalamus_gene.signf.pvals)
colnames(pituitary.vs.hypothalamus_gene.signf.pvals)<- c("p_value")

pituitary.vs.hypothalamus_gene.signf.pvals$geneID <- row.names(pituitary.vs.hypothalamus_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than 0.01
filtered_pituitary.vs.hypothalamus_gene.signf.pvals <- pituitary.vs.hypothalamus_gene.signf.pvals[pituitary.vs.hypothalamus_gene.signf.pvals$p_value <= adjusted_p_value,]
pituitary.vs.hypothalamus_gene.signf.pvals <- na.omit(filtered_pituitary.vs.hypothalamus_gene.signf.pvals)

# Print the filtered data frame
print(filtered_pituitary.vs.hypothalamus_gene.signf.pvals)
```


# Signficantly correlated genes pituitary.vs.pwbc : 
```{r}
pituitary.vs.pwbc_gene.signf.pvals<-data.frame(pituitary.vs.pwbc_gene.signf.pvals)
colnames(pituitary.vs.pwbc_gene.signf.pvals)<- c("p_value")

pituitary.vs.pwbc_gene.signf.pvals$geneID <- row.names(pituitary.vs.pwbc_gene.signf.pvals)

# Filter out all geneID that are associated to a p value lower than 0.01
filtered_pituitary.vs.pwbc_gene.signf.pvals <- pituitary.vs.pwbc_gene.signf.pvals[pituitary.vs.pwbc_gene.signf.pvals$p_value <= adjusted_p_value,]
pituitary.vs.pwbc_gene.signf.pvals <- na.omit(filtered_pituitary.vs.pwbc_gene.signf.pvals)

# Print the filtered data frame
print(filtered_pituitary.vs.pwbc_gene.signf.pvals)
```

```{r}
# Combine the geneID columns into a single vector
all_de.gene_ids <- c(
  pituitary.vs.hypothalamus_gene.signf.pvals$geneID,
  pwbc.vs.hypothalamus_gene.signf.pvals$geneID,
  pituitary.vs.endometrium_gene.signf.pvals$geneID,
  pwbc.vs.endometrium_gene.signf.pvals$geneID,
  hypothalamus.vs.endometrium_gene.signf.pvals$geneID,
  pituitary.vs.pwbc_gene.signf.pvals$geneID
)
```

```{r}
all_de.gene_ids
```

```{r}
# Remove duplicates from the geneID vector
unique_de.gene_ids <- unique(all_de.gene_ids)
```

```{r}
measured_genes<-colnames(CTF_normalized_asinh_)
```

```{r}
length(unique_de.gene_ids)
