# Load libraries 

```{r}
library('dplyr', quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library('ggplot2', quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library('edgeR',quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("cowplot", quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library("GGally",quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library("DESeq2",quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library('readxl',quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library("VennDiagram", quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("ComplexHeatmap",quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("flashClust", quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library('plotly' , quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library('tidyverse' ,quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library('htmlwidgets' , quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library('reshape2' ,quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("ggpubr" ,quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("car" ,quietly = TRUE, lib.loc="/usr/lib/R/site-library") 
library("goseq" ,quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("stringr",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("data.table",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("tidyr",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("dplyr",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("ggsignif",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("kableExtra",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("grid",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("gridExtra",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("bigmemory",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("doParallel",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("ggmanh",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("lme4",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("emmeans",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("limma",quietly = TRUE , lib.loc="/usr/lib/R/site-library") 
library("MultiAssayExperiment",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library('MOFA2', lib.loc="/usr/lib/R/site-library")
library(clusterProfiler)
library(enrichplot)
library(readxl)
library(ggridges)
organism = "org.Bt.eg.db"
library(organism, character.only = TRUE)
```
# Load the genes that are signficantly co-expressed with their associated p-value

```{r}
# paths need to be adapted 
hypothalamus.vs.endometrium_gene_list<-read_excel("/home/yassine/Downloads/hypothalamus.vs.endometrium.xlsx")
pituitary.vs.endometrium_gene_list<-read_excel("/home/yassine/Downloads/pituitary.vs.endometrium.xlsx")
pituitary.vs.hypothalamus_gene_list<-read_excel("/home/yassine/Downloads/pituitary.vs.hypothalamus.xlsx")
pituitary.vs.pwbc_gene_list<-read_excel("/home/yassine/Downloads/pituitary.vs.pwbc.xlsx")
pwbc.vs.hypothalamus_gene_list<-read_excel("/home/yassine/Downloads/pwbc.vs.hypothalamus.xlsx")
pwbc.vs.endometrium_gene_list<-read_excel("/home/yassine/Downloads/pwbc.vs.endometrium.xlsx")

```
# Wrangling the gene lists to the appropriate clusterProfilier input format

```{r}
# we want the p values  
original_gene_list <- hypothalamus.vs.endometrium_gene_list$p_value
# name the vector
names(original_gene_list) <- hypothalamus.vs.endometrium_gene_list$geneID
# omit any NA values 
hypothalamus.vs.endometrium_list<-na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
hypothalamus.vs.endometrium_list = sort(hypothalamus.vs.endometrium_list, 
                                        decreasing = TRUE)


original_gene_list <- pituitary.vs.endometrium_gene_list$p_value

names(original_gene_list) <- pituitary.vs.endometrium_gene_list$geneID

pituitary.vs.endometrium_gene_list<-na.omit(original_gene_list)

pituitary.vs.endometrium_gene_list = sort(pituitary.vs.endometrium_gene_list, 
                                        decreasing = TRUE)


original_gene_list <- pituitary.vs.hypothalamus_gene_list$p_value

names(original_gene_list) <- pituitary.vs.hypothalamus_gene_list$geneID

pituitary.vs.hypothalamus_gene_list<-na.omit(original_gene_list)

pituitary.vs.hypothalamus_gene_list = sort(pituitary.vs.hypothalamus_gene_list, 
                                        decreasing = TRUE)


original_gene_list <- pituitary.vs.pwbc_gene_list$p_value

names(original_gene_list) <- pituitary.vs.pwbc_gene_list$geneID

pituitary.vs.pwbc_gene_list<-na.omit(original_gene_list)

pituitary.vs.pwbc_gene_list = sort(pituitary.vs.pwbc_gene_list, 
                                        decreasing = TRUE)

original_gene_list <- pwbc.vs.hypothalamus_gene_list$p_value

names(original_gene_list) <- pwbc.vs.hypothalamus_gene_list$geneID

pwbc.vs.hypothalamus_gene_list<-na.omit(original_gene_list)

pwbc.vs.hypothalamus_gene_list = sort(pwbc.vs.hypothalamus_gene_list, 
                                        decreasing = TRUE)

original_gene_list <- pwbc.vs.endometrium_gene_list$p_value

names(original_gene_list) <- pwbc.vs.endometrium_gene_list$geneID

pwbc.vs.endometrium_gene_list<-na.omit(original_gene_list)

pwbc.vs.endometrium_gene_list = sort(pwbc.vs.endometrium_gene_list, 
                                        decreasing = TRUE)


# inspecting the gene lists 

```{r}
#print(hypothalamus.vs.endometrium_list)

#print(pituitary.vs.endometrium_gene_list)

#print(pituitary.vs.hypothalamus_gene_list)

#print(pwbc.vs.hypothalamus_gene_list)

#print(pituitary.vs.pwbc_gene_list)

#print(pwbc.vs.endometrium_gene_list)
```
# Pperforming the GO using  gseGO() function from clusterProfiler

```{r}
gse_hypothalamus.vs.endometrium <- gseGO(geneList=hypothalamus.vs.endometrium_list, 
             ont ="BP", 
             keyType = "ENSEMBL", 
             nPerm = 10e6, 
             minGSSize = 5, 
             maxGSSize = 1000, 
             pvalueCutoff = 0.01, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")

gse_pituitary.vs.endometrium <- gseGO(geneList=pituitary.vs.endometrium_gene_list, 
             ont ="BP", 
             keyType = "ENSEMBL", 
             nPerm = 10e6, 
             minGSSize = 5, 
             maxGSSize = 1000, 
             pvalueCutoff = 0.01, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")

gse_pituitary.vs.hypothalamus <- gseGO(geneList=pituitary.vs.hypothalamus_gene_list, 
             ont ="BP", 
             keyType = "ENSEMBL", 
             nPerm = 10e6, 
             minGSSize = 5, 
             maxGSSize = 1000, 
             pvalueCutoff = 0.01, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")

gse_pwbc.vs.hypothalamus <- gseGO(geneList=pwbc.vs.hypothalamus_gene_list, 
             ont ="BP", 
             keyType = "ENSEMBL", 
             nPerm = 10e6, 
             minGSSize = 5, 
             maxGSSize = 1000, 
             pvalueCutoff = 0.01, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")


gse_pituitary.vs.pwbc <- gseGO(geneList=pituitary.vs.pwbc_gene_list, 
             ont ="BP", 
             keyType = "ENSEMBL", 
             nPerm = 10e6, 
             minGSSize = 5, 
             maxGSSize = 1000, 
             pvalueCutoff = 0.01, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")

gse_pwbc.vs.endometrium <- gseGO(geneList=pwbc.vs.endometrium_gene_list, 
             ont ="BP", 
             keyType = "ENSEMBL", 
             nPerm = 10e6, 
             minGSSize = 5, 
             maxGSSize = 1000, 
             pvalueCutoff = 0.01, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")
```

# Visualising the GO results 

# 1.dotplot

```{r}
require(DOSE)

dotplot_hypothalamus.vs.endometrium<-dotplot(gse_hypothalamus.vs.endometrium,
                                             showCategory=10,
                                             split=".sign") + 
  facet_grid(.~.sign)+ 
  labs(title = "Enriched BPs between hypothalamus and endometrium tissues ",
       subtitle = "activated = co-expressed by both tissues , suppressed = opposite expression pattern in both tissues ")
```

```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/dotplot_hypothalamus.vs.endometrium.png", # path need to be adapted
       plot = dotplot_hypothalamus.vs.endometrium, 
       width = 10, 
       height = 10)

```

```{r}
dotplot_pwbc.vs.endometrium<-dotplot(gse_pwbc.vs.endometrium,
                                     showCategory=10,
                                     split=".sign") + 
  facet_grid(.~.sign)+ 
  labs(title = "Enriched BPs between pwbc and endometrium tissues ",
       subtitle = "activated = co-expressed by both tissues , suppressed = opposite expression pattern in both tissues ")
```

```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/dotplot_pwbc.vs.endometrium.png", # path need to be adapted
       plot = dotplot_pwbc.vs.endometrium, 
       width = 10, 
       height = 10)
```

```{r}

dotplot_pituitary.vs.endometrium<-dotplot(gse_pituitary.vs.endometrium,
        showCategory=10,
        split=".sign") + 
  facet_grid(.~.sign)+ 
  labs(title = "Enriched BPs between pituitary and endometrium tissues ",
       subtitle = "activated = co-expressed by both tissues , suppressed = opposite expression pattern in both tissues ")
```

```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/dotplot_pituitary.vs.endometrium.png", # path need to be adapted
       plot = dotplot_pituitary.vs.endometrium, 
       width = 10, 
       height = 10)
```

```{r}
dotplot_pituitary.vs.hypothalamus <-dotplot(gse_pituitary.vs.hypothalamus,
        showCategory=10,
        split=".sign") + 
  facet_grid(.~.sign)+ 
  labs(title = "Enriched BPs between pituitary and hypothalamus tissues ",
       subtitle = "activated = co-expressed by both tissues , suppressed = opposite expression pattern in both tissues ")
```

```{r}

ggsave(filename = "/home/yassine/Downloads/Figures/dotplot_pituitary.vs.hypothalamus.png", # path need to be adapted
       plot = dotplot_pituitary.vs.hypothalamus, 
       width = 10, 
       height = 10)
```

```{r}
dotplot_pwbc.vs.hypothalamus<-dotplot(gse_pwbc.vs.hypothalamus,
        showCategory=10,
        split=".sign") + 
  facet_grid(.~.sign)+
  labs(title = "Enriched BPs between pwbc and hypothalamus tissues ",
       subtitle = "activated = co-expressed by both tissues , suppressed = opposite expression pattern in both tissues ")

```

```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/dotplot_pwbc.vs.hypothalamus.png", # path need to be adapted
       plot = dotplot_pwbc.vs.hypothalamus, 
       width = 10, 
       height = 10)
```

```{r}
dotplot_pituitary.vs.pwbc<-dotplot(gse_pituitary.vs.pwbc,
        showCategory=10,
        split=".sign") + 
  facet_grid(.~.sign)+ 
  labs(title = "Enriched BPs between pituitary and pwbc tissues ",
       subtitle = "activated = co-expressed by both tissues , suppressed = opposite expression pattern in both tissues ")

```

```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/dotplot_pituitary.vs.pwbc.png", # path need to be adapted
       plot = dotplot_pituitary.vs.pwbc, 
       width = 10, 
       height = 10)
```

# 2.Category Netplot

```{r}
cnetplot_hypothalamus.vs.endometrium<-cnetplot(gse_hypothalamus.vs.endometrium, 
         categorySize="pvalue",
         p=hypothalamus.vs.endometrium_list, 
         showCategory = 10,
         layout = "kk")

```

```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/cnetplot_hypothalamus.vs.endometrium.png", # path need to be adapted
       plot = cnetplot_hypothalamus.vs.endometrium, 
       width = 10, 
       height = 10)
```

```{r}
cnetplot_pituitary.vs.endometrium<-cnetplot(gse_pituitary.vs.endometrium, 
         categorySize="pvalue",
         p=pituitary.vs.endometrium_list, 
         showCategory = 10,
         layout = "kk")
```

```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/cnetplot_pituitary.vs.endometrium.png", # path need to be adapted
       plot = cnetplot_pituitary.vs.endometrium, 
       width = 10, 
       height = 10)
```
```{r}
cnetplot_pituitary.vs.hypothalamus<-cnetplot(gse_pituitary.vs.hypothalamus, 
         categorySize="pvalue",
         p=pituitary.vs.hypothalamus_list, 
         showCategory = 10,
         layout = "kk")
```

```{r}

ggsave(filename = "/home/yassine/Downloads/Figures/cnetplot_pituitary.vs.hypothalamus.png", # path need to be adapted
       plot = cnetplot_pituitary.vs.hypothalamus, 
       width = 10, 
       height = 10)
```

```{r}
cnetplot_pituitary.vs.pwbc<-cnetplot(gse_pituitary.vs.pwbc, 
         categorySize="pvalue",
         p=pituitary.vs.pwbc_list, 
         showCategory = 10,
         layout = "kk")
```
```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/cnetplot_pituitary.vs.pwbc.png", # path need to be adapted
       plot = cnetplot_pituitary.vs.pwbc, 
       width = 10, 
       height = 10)
```

```{r}
cnetplot_pwbc.vs.endometrium<-cnetplot(gse_pwbc.vs.endometrium, 
         categorySize="pvalue",
         p=pwbc.vs.endometrium_list, 
         showCategory = 10,
         layout = "kk")
```

```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/cnetplot_cnetplot_pwbc.vs.endometrium.png", # path need to be adapted
       plot = cnetplot_pwbc.vs.endometrium, 
       width = 10, 
       height = 10)
```

```{r}
cnetplot_pwbc.vs.hypothalamus<-cnetplot(gse_pwbc.vs.hypothalamus, 
         categorySize="pvalue",
         p=pwbc.vs.hypothalamus_list, 
         showCategory = 10,
         layout = "kk")
```


```{r}
ggsave(filename = "/home/yassine/Downloads/Figures/cnetplot_pwbc.vs.hypothalamus.png", # path need to be adapted
       plot = cnetplot_pwbc.vs.hypothalamus, 
       width = 10, 
       height = 10)
```
