```{r}
# For data management
BiocManager::install("org.Hs.eg.db")
# For visualisation
install.packages('pheatmap')
install.packages("DOSE")
install.packages("enrichplot")
install.packages("ggupset")

```

# Length data and category mapping :

```{r}
annotation.ensembl.symbol<- read.table("/home/yassine/Downloads/2023_10_10_btaurus_annotation.ensembl.symbol.txt",
                                       header = TRUE,
                                       sep= "\t")

annotation.ensembl.transcript<-read.table("/home/yassine/Downloads/2023_10_10_btaurus_annotation.ensembl.transcript.txt",
                                       header = TRUE,
                                       sep= "\t")

annotation.GO.biomart<- read.table("/home/yassine/Downloads/2023_10_10_btaurus_annotation.GO.biomart.txt", 
                                   header = TRUE,
                                   sep= "\t")
gene.length<-read.table("/home/yassine/Downloads/2023_10_10_btaurus_gene.length.txt", 
                        header = TRUE,
                        sep= "\t")

```


```{r}
CTF_asinh_count_data_<-read.table("/home/yassine/Downloads/CTF_asinh_count_data.txt", 
                        header = TRUE,
                        sep= "\t")
measured_genes<-colnames(CTF_asinh_count_data_)
```

```{r}
all_genes<-data.frame(gene=measured_genes,
                      stringsAsFactors=FALSE )
rownames(all_genes)<-all_genes$gene
N_expressed_genes<-length(all_genes$gene)
gene.length_<-gene.length[gene.length$ensembl_gene_id %in% all_genes$gene,]
annotation.genelength.biomart_vector<-gene.length_$transcript_length
annotation.genelength.biomart_vector<-data.frame(annotation.genelength.biomart_vector)
row.names(annotation.genelength.biomart_vector)<-gene.length_$ensembl_gene_id
```


```{r}
annotation.GO.BP.biomart<-annotation.GO.biomart[annotation.GO.biomart$namespace_1003=="biological_process", c(1,3)]

annotation.GO.BP.biomart<-annotation.GO.BP.biomart[annotation.GO.BP.biomart$ensembl_gene_id %in% rownames(all_genes),]

annotation.GO.MF.biomart<-annotation.GO.biomart[annotation.GO.biomart$namespace_1003=="molecular_function", c(1,3)] 

annotation.GO.MF.biomart<-annotation.GO.MF.biomart[annotation.GO.MF.biomart$ensembl_gene_id %in% rownames(all_genes),]
```

# The different genes to test :  
# pituitary.vs.hypothalamus_gene.signf.pvals$geneID,
# pwbc.vs.hypothalamus_gene.signf.pvals$geneID,
# pituitary.vs.endometrium_gene.signf.pvals$geneID,
# pwbc.vs.endometrium_gene.signf.pvals$geneID,
# hypothalamus.vs.endometrium_gene.signf.pvals$geneID,
# pituitary.vs.pwbc_gene.signf.pvals$geneID
```{r}
pituitary.vs.hypothalamus_gene.signf.pvals <-read.table("/home/yassine/Downloads/pituitary.vs.hypothalamus_signf.gene.txt",
                                                        header = TRUE,
                                                        sep ="\t" )
pwbc.vs.hypothalamus_gene.signf.pvals<-read.table("/home/yassine/Downloads/pwbc.vs.hypothalamus_signf.gene.txt",
                                                        header = TRUE,
                                                        sep ="\t" )
pituitary.vs.endometrium_gene.signf.pvals<-read.table("/home/yassine/Downloads/pituitary.vs.endometrium_signf.gene.txt",
                                                        header = TRUE,
                                                        sep ="\t" )
pwbc.vs.endometrium_gene.signf.pvals<-read.table("/home/yassine/Downloads/pwbc.vs.endometrium_signf.gene.txt",
                                                        header = TRUE,
                                                        sep ="\t" )
hypothalamus.vs.endometrium_gene.signf.pvals<-read.table("/home/yassine/Downloads/hypothalamus.vs.endometrium_signf.gene.txt",
                                                        header = TRUE,
                                                        sep ="\t" )
pituitary.vs.pwbc_gene.signf.pvals<-read.table("/home/yassine/Downloads/pituitary.vs.pwbc_signf.gene.txt",
                                                        header = TRUE,
                                                        sep ="\t" )

```


```{r}
test.genes<-data.frame(a=pituitary.vs.hypothalamus_gene.signf.pvals$geneID,
                       stringsAsFactors=FALSE)
```


```{r}
all_genes_numeric<-as.integer(all_genes$gene %in% test.genes$a)
all_genes_numeric<-data.frame(all_genes_numeric)
row.names(all_genes_numeric)<-all_genes$gene
N_sig_genes<-length(test.genes$a)
colnames(annotation.genelength.biomart_vector)= "transcript_length"
```


```{r}
#N_sig_genes
set.seed(9830)
pwf<-nullp(DEgenes=all_genes_numeric$all_genes_numeric,
           bias.data=annotation.genelength.biomart_vector$transcript_length,
           plot.fit=FALSE )
```

```{r}
go_map=getgo(rownames(all_genes),
             "bosTau4",
             "ensGene")

```

# Stopped here , couldn't debug the error 

```{r}
GO_BP_Cats_raw_counts<-goseq(pwf= pwf,
                             gene2cat= annotation.GO.BP.biomart ,#go_map,
                             method ="Sampling",
                             repcnt = 5000,
                             use_genes_without_cat= FALSE)
```


```{r}
GO_BP_Cats_raw_counts<-GO_BP_Cats_raw_counts[GO_BP_Cats_raw_counts$numDEInCat>3,]
GO_BP_Cats_raw_counts$FWER<-p.adjust(GO_BP_Cats_raw_counts$over_represented_pvalue,
                                     method ="holm")

GO_BP_Cats_raw_counts<-GO_BP_Cats_raw_counts[with(GO_BP_Cats_raw_counts,
                                                  order(FWER,over_represented_pvalue,
                                                        -numDEInCat)), ]
```

```{r}
#head(GO_BP_Cats_raw_counts, n=20)
```

```{r}
GO_BP_Cats_raw_counts$fold_enrichment<-(GO_BP_Cats_raw_counts$numDEInCat/N_sig_genes)/(G O_BP_Cats_raw_counts$numInCat/N_expressed_genes) annotation.GO.BP.biomart_testgenes<-annotation.GO.BP.biomart[annotation.GO.BP.biomart$en sembl_gene_id %in% test.genes$a, ] GO_BP_Cats_raw_counts<-merge(GO_BP_Cats_raw_counts,annotation.GO.BP.biomart_testgenes, b y.x="category", by.y="go_id", all.x=TRUE, all.y=FALSE) GO_BP_Cats_raw_counts<-merge(GO_BP_Cats_raw_counts, annotation.ensembl.symbol, by.x="ens embl_gene_id", by.y="ensembl_gene_id", all=FALSE, all.x=TRUE, all.y=FALSE) GO_BP_Cats_raw_counts<-GO_BP_Cats_raw_counts[with(GO_BP_Cats_raw_counts, order(FWER,ter m)), ]
    GO_BP_Cats_raw_counts<-GO_BP_Cats_raw_counts[GO_BP_Cats_raw_counts$FWER<0.1,]
