# Load the needed libraries

```{r}
library('dplyr', quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library('ggplot2', quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library('edgeR',quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("cowplot", quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library("GGally",quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library("DESeq2",quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library('readxl',quietly = TRUE,lib.loc="/usr/lib/R/site-library")
library("VennDiagram", quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("ComplexHeatmap",quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("flashClust", quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library('plotly' , quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library('tidyverse' ,quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library('htmlwidgets' , quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library('reshape2' ,quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("ggpubr" ,quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("car" ,quietly = TRUE, lib.loc="/usr/lib/R/site-library") 
library("goseq" ,quietly = TRUE, lib.loc="/usr/lib/R/site-library")
library("stringr",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("data.table",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("tidyr",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("dplyr",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("ggsignif",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("kableExtra",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("grid",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("gridExtra",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("bigmemory",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("doParallel",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("ggmanh",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("lme4",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("emmeans",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library("limma",quietly = TRUE , lib.loc="/usr/lib/R/site-library") 
library("MultiAssayExperiment",quietly = TRUE , lib.loc="/usr/lib/R/site-library")
library('MOFA2', lib.loc="/usr/lib/R/site-library")
```

# Load the counts.summary files 

```{r}
# change the path to the folder that contains the counts.summary folder

"/mnt/storage/lab_folder/counting/"
files_<-list.files("/mnt/storage/lab_folder/counting/",
                  recursive=T,
                  pattern="summary",
                  full.names = TRUE)
```


```{r}
# check the number of files loaded 

length(files_)
```

# The counts of the reads belonging to each category per sample

```{r}
# Combines the content of the counts.summary files into a single table showing the number of 
# reads belonging to each category per sample 
 
number_files <- length(files_)

reads_produced_a<-data.frame()
reads_produced_b<-data.frame()

for (n in 1:number_files) {
reads_produced<-read.delim(files_[n],
                           sep= "\t",
                           header=FALSE, 
                           stringsAsFactors = FALSE, 
                           comment.char= "#")
sample<-substring(files_[n], 54,56)  # Needs to be changed to get the numbers or the single identifiers of each of your samples 
number_of_reads_produced <- 0
  for (k in 2:15) {
    number_of_reads_produced = number_of_reads_produced + as.integer(reads_produced[k,2])
  }
reads_produced_a<-data.frame(sample,number_of_reads_produced)
reads_produced_b<-rbind(reads_produced_b,reads_produced_a)
}
```

# Builing the counting_summary_b table (see READ_ME for more informations about it's content)

```{r}
counting_summary_a<-data.frame()
counting_summary_b<-data.frame()

for (n in 1:number_files) {
counting_summary<-read.delim(files_[n],
                             sep= "\t",
                             header=FALSE,
                             stringsAsFactors = FALSE,
                             comment.char= "#")
sample<-substring(files_[n], 54,56) # Needs to be changed to get the numbers or the single identifiers of each of your samples

number_of_reads_assigned<-as.integer(counting_summary[2,2])
unassigned_NoFeatures<-as.integer(counting_summary[13,2])
unassigned_Ambiguity<-as.integer(counting_summary[15,2])

counting_summary_a<-data.frame(sample,
                               number_of_reads_assigned, 
                               unassigned_NoFeatures, 
                               unassigned_Ambiguity)
counting_summary_b<-rbind(counting_summary_b,
                          counting_summary_a)
}

counting_summary_b$reads_sequenced <-reads_produced_b$number_of_reads_produced

counting_summary_b$reads_retained <- counting_summary_b$number_of_reads_assigned + counting_summary_b$unassigned_NoFeatures + counting_summary_b$unassigned_Ambiguity

counting_summary_b$reads_discarted<-counting_summary_b$reads_sequenced - counting_summary_b$reads_retained

counting_summary_b$perc_number_of_reads_discarted<-counting_summary_b$reads_discarted/counting_summary_b$reads_sequenced

counting_summary_b$perc_number_of_reads_retained<-counting_summary_b$reads_retained/counting_summary_b$reads_sequenced

counting_summary_b$perc_number_of_reads_assigned<-counting_summary_b$number_of_reads_assigned/counting_summary_b$reads_sequenced

counting_summary_b$perc_unassigned_NoFeatures<-counting_summary_b$unassigned_NoFeatures/counting_summary_b$reads_sequenced

counting_summary_b$perc_unassigned_Ambiguity<-counting_summary_b$unassigned_Ambiguity/counting_summary_b$reads_sequenced
```

# Identifying samples with number of produced reads lower than thresh

```{r}

thresh <-1e+6 # can be changed 

samples_with_low_reads <- counting_summary_b[counting_summary_b$reads_sequenced < thresh,]

# Get the sample names
sample_names <- samples_with_low_reads$sample

# Print the sample names
print(sample_names)

```
# Check the samples with low read couts 

```{r}
samples_with_low_reads
```

# Filter out the samples with number of produced reads lower than thresh

```{r}
samples_with_high_reads <- counting_summary_b[counting_summary_b$reads_sequenced > 1e+6,]
```

# Summary statistics
```{r}

# Get the percentages and absolut counts of Unassigned_Ambiguity, Unassigned_NoFeatures, and Assigned reads for samples that have a number of assigned reads larger that 1e+6

percentages_of_unassigned_ambiguity <- samples_with_high_reads$perc_unassigned_Ambiguity
percentages_of_unassigned_no_features <- samples_with_high_reads$perc_unassigned_NoFeatures
percentages_of_assigned_reads <- samples_with_high_reads$perc_number_of_reads_assigned

number_unassigned_ambiguity <- samples_with_high_reads$unassigned_Ambiguity
number_of_unassigned_no_features <- samples_with_high_reads$unassigned_NoFeatures
number_of_assigned_reads <- samples_with_high_reads$number_of_reads_assigned

# Calculate the mean and variance across samples for both the percentages and absolut numbers

mean_percentages_of_unassigned_ambiguity <- mean(percentages_of_unassigned_ambiguity)
mean_percentages_of_unassigned_no_features <- mean(percentages_of_unassigned_no_features)
mean_percentages_of_assigned_reads <- mean(percentages_of_assigned_reads)

mean_number_of_unassigned_ambiguity <- mean(number_unassigned_ambiguity)
mean_number_of_unassigned_no_features <- mean(number_of_unassigned_no_features)
mean_number_of_assigned_reads <- mean(number_of_assigned_reads)


variance_percentages_of_unassigned_ambiguity <- var(percentages_of_unassigned_ambiguity)
variance_percentages_of_unassigned_no_features <- var(percentages_of_unassigned_no_features)
variance_percentages_of_percentages_of_assigned_reads <- var(percentages_of_assigned_reads)

variance_number_of_unassigned_ambiguity <- var(number_unassigned_ambiguity)
variance_number_of_unassigned_no_features <- var(number_of_unassigned_no_features)
variance_number_of_assigned_reads <- var(number_of_assigned_reads)

```

```{r}
summary_percentages <- data.frame(mean = c(mean_percentages_of_unassigned_ambiguity,
                                           mean_percentages_of_unassigned_no_features,
                                           mean_percentages_of_assigned_reads),
                                 variance = c(variance_percentages_of_unassigned_ambiguity,
                                              variance_percentages_of_unassigned_no_features,
                                          variance_percentages_of_percentages_of_assigned_reads))

colnames(summary_percentages) <- c("mean", "variance")
rownames(summary_percentages) <- c("perc_unassigned_ambiguity",
                                   "perc_unassigned_no_features",
                                   "perc_assigned_reads")

summary_numbers <- data.frame(mean = c(mean_number_of_unassigned_ambiguity,
                                           mean_number_of_unassigned_no_features,
                                           mean_number_of_assigned_reads),
                                 variance = c(variance_number_of_unassigned_ambiguity,
                                              variance_number_of_unassigned_no_features,
                                          variance_number_of_assigned_reads))

colnames(summary_numbers) <- c("mean", "variance")
rownames(summary_numbers) <- c("unassigned_ambiguity",
                              "unassigned_no_features",
                               "assigned_reads")
```

# Print the summary tables

```{r}
print(summary_percentages)

print(summary_numbers)
```

# Creates a dataframe with perc_number_of_reads_retained / perc_number_of_reads_assigned / perc_unassigned_NoFeatures and perc_unassigned_Ambiguity needed for plot a 

```{r}
counting_summary_c<-reshape2::melt(counting_summary_b[,c(1,9,10,11,12)])

counting_summary_c$value<-counting_summary_c$value*100
counting_summary_c$variable<-factor(counting_summary_c$variable,
                                    levels=c( "perc_number_of_reads_retained" ,
                                              "perc_number_of_reads_assigned",
                                              "perc_unassigned_NoFeatures" ,
                                              "perc_unassigned_Ambiguity"))
```
# plot_a shows the perc_number_of_reads_retained / perc_number_of_reads_assigned / perc_unassigned_NoFeatures and perc_unassigned_Ambiguity for each sample 

```{r}
plot_a<-ggplot() + 
  geom_bar(aes(y = value,
               x = sample, 
               fill = variable),
           data = counting_summary_c,
           stat="identity")+
  scale_y_continuous(name="Percentage", 
                     breaks = seq(0,5,105))+
  scale_fill_hue(labels=c("Retained", 
                          "Assigned",
                          "Unassigned_NoFeatures" ,
                          "Unassigned_Ambiguity"))+
  #ggtitle("Distribution of reads after alignment")+
  theme_bw(base_size = 12)+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle=75, hjust=1))
```
# plot_b shows the absolut number of sequenced reads for each sample 

```{r}
counting_summary_d<-reshape2::melt(counting_summary_b[,c(1,5)])
plot_b<-ggplot() + 
  geom_bar(aes(y = value,
               x = sample, 
               fill = variable),
           data = counting_summary_d,stat="identity")+
  scale_y_continuous(name="Read pairs sequenced")+
  theme_bw(base_size = 12)+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank())
```
# plot_c shows the absolut number of reads assigned for each sample

```{r}
counting_summary_e<-reshape2::melt(counting_summary_b[,c(1,2)])
plot_c<-ggplot() + 
  geom_bar(aes(y = value,
               x = sample, 
               fill = variable),
           data = counting_summary_e,stat="identity")+
  scale_y_continuous(name="Read pairs in annotation")+
  geom_hline(yintercept=10^6,
             linetype="dashed",
             color = "gray",
             size=1)+
  theme_bw(base_size = 12)+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank())
```
# plot_d shows the percentage of reads assigned for each sample

```{r}
counting_summary_f<-reshape2::melt(counting_summary_b[,c(1,10)])
plot_d<-ggplot() + 
  geom_bar(aes(y = value,
               x = sample, 
               fill = variable), 
           data = counting_summary_f,stat="identity")+
  scale_y_continuous(name="Proportion reads \n matching annotation")+
  theme_bw(base_size = 12)+
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank())
```

# Combine the previously created plots in a grid 

```{r}
grid<-plot_grid( plot_b,
           plot_c,
           plot_d,
           plot_a,  
           ncol = 1, 
           align = 'v')
grid
```
